WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2019-03-27 review/referee_response_0.txt

The authors present a number of new and high-quality transit observations of
the hot Jupiter WASP-4b observed by TESS. After comparing the predicted
mid-transit times (obtained only considering the pre-existing transits in the
literature) with the measured values from the TESS light curves, the authors
conclude that the TESS transits occur ~ 82 seconds earlier than predicted. They
examine different possibilities and find that the 2 most favored scenarios to
explain this deviation are orbital decay and apsidal precession. However, the
authors caution that more observations are needed to confirm these results.
Also, they redetermine various parameters of the system, finding them in
general agreement with literature values.

In general, the manuscript is scientifically accurate and the results presented
in it certainly will be of interest to other investigators in the field. Then,
I consider this paper worthy of being published after the concerns described
below be addressed.

Below are my comments, beginning with the major points. I recommend the authors
take them into consideration when revising their manuscript.

  > RESPONSE: We thank the referee for this summary. The major and minor
  > points are addressed individually below. There are also three larger
  > changes that we highlight here.
  >
  > 1) In our revision, we have added weight to a third possible
  > explanation that we previously mentioned but did not sufficiently
  > emphasize: a massive outer companion, perhaps on an eccentric orbit,
  > might be able to explain both the transit and radial velocity
  > observations. We have described this possibility throughout the text
  > where appropriate.
  >
  > 2) We also added our own re-analysis of the extant RV data (now
  > section 4.3, and figure 6) to check on the Knutson et al (2014)
  > result that the star's acceleration towards our line of sight is
  > only weakly detected, and is an order of magnitude too small to
  > explain the observed period derivative from the transit times.
  >
  > 3) Finally, we have also added a more detailed exploration of how
  > the assumed prior in the precession model affects the inferred Love
  > number, and the predicted transit and occultation times.

@@

REFEREE COMMENT

2) As mentioned in 1), another possible interpretation for explaining the
deviation from a linear ephemeris of the TESS mid-transit times could be the
presence of anomalies in the light curves due to stellar activity. It has been
well established that WASP-4 is an active star with a rotation period (P) in
between 20-40 days (P ~ 22 days and P ~ 34 days were determined by
Sanchis-Ojeda et al. 2011 and Hoyer et al. 2013, respectively), which is
expected for a main-sequence G7 star. Given that the TESS mid-transits span
around 30 days, wouldn't it be possible that these measurements be affected by
the existence of unseen stellar spots? Regarding this point, I would like to
see an extra paragraph in the Section "Interpretation" with a discussion about
this scenario.

  > RESPONSE: We have considered this possibility, and have now added
  > the following text in Section 3.3:

  > To explore the effect of possible spot-crossing events on the TESS
  > transit time measurements, we performed a separate test.  We
  > injected triangular spot-anomalies with amplitude $0.03\%$ and
  > duration 30 minutes at random phases into each transit.  The
  > amplitudes were larger than the spot-crossing anomalies observed by
  > \citet{southworth_high-precision_2009} and
  > \citet{sanchis-ojeda_starspots_2011}, and the durations were
  > comparable.  Spots of these amplitudes resemble the anomalies
  > present in transits ``1360.54'' and ``1372.58'' of Figure
  > \ref{fig:lightcurves}: they are visible, but at low confidence.
  > 
  > With spots injected, we repeated our measurement of the transit
  > times.  On average, the measured transit times did not change after
  > injecting spots, because the flux deviations are equally likely to
  > occur in the first and second halves of the transit.  For individual
  > transits, there were no cases for which the timing deviation was
  > larger than one minute.  The largest shifts occur when the spot
  > anomaly occurs during transit ingress or egress, in which case the
  > measured mid-time is shifted either late or early by between 30 and
  > 50 seconds \citep[qualitatively similar to results found
  > by][]{ioannidis_how_2016}.
  > 
  > Therefore the TESS observations could hypothetically all be skewed
  > early if there were spot-crossing events during every egress.  Two
  > arguments rule out this possibility.  (1) The lightcurve residuals
  > do not show evidence for these events.  (2) The stellar rotation
  > period is between 20 and 40 days, and the sky-projected stellar
  > obliquity is less than 10 degrees
  > \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
  > Since the planet orbits every 1.3 days, requiring that spot
  > anomalies always occur during egress would be equivalent to
  > requiring a stellar spot distribution that is exquisitely (and thus
  > implausibly) distributed to match the planet egress times.

@@

  > RESPONSE: Done.

REFEREE COMMENT

iii) For the casual reader, it would be important to briefly specify what
"Threshold Crossing Events" means, and what the flags "Reaction Wheel
Desaturation Event" and "Manual Exclude" represent. Footnotes including this
information would be appropriate. Also, could the authors describe what a
"Presearch Data Conditioning" lightcurve is?

  > RESPONSE: We have clarified that a "Presearch Data Conditioning"
  > lightcurve is at a stage of process in which systematic trends have
  > been removed through the decorrelation process described by Smith et
  > al (2017).
  >
  > We now omit explicit mention of "Treshold Crossing Events" in favor
  > of a more qualitative description.
  >
  > As described in the text, the flags "Reaction Wheel Desaturation
  > Event" and "Manual Exclude" together mean that the data are
  > potentially of poor quality due to "momentum dumps".  We referred
  > the reader to the data release notes for an explanation of what a
  > "momentum dump" is, and have added text to this footnote to improve
  > its clarity.

@@

REFEREE COMMENT

iv) It might be worthwhile to mention what "ramp-like systematic effects" are
and also include a reference.

  > RESPONSE: We have changed the wording from "ramp-like systematic
  > effects" to "correlated red noise". No reference is needed.

@@

  > RESPONSE: Done.

REFEREE COMMENT

v) Similar to i) How did you fit the mid-transit times and occultations shown
in Figure 4?

  > RESPONSE: We have added the following sentences in Section 3.2:
  > """
  > We fitted each model by assuming a Gaussian likelihood and sampling
  > over the posterior probability distributions. The prior for the
  > quadratic model allowed the period derivative to have any sign.  We
  > considered two possible priors for the precession model: the first
  > is a wide prior that allows non-physical values of the planetary
  > Love number (Equation~\ref{eq:love_number}). The second requires
  > that the planetary Love number is less than that of a sphere with a
  > homogeneous density profile.  
  > """

@@

  > RESPONSE: Changed wording to remove any ambiguity.

REFEREE COMMENT:

iv) The authors mention that there are about 20 Hot Jupiters for which the
theoretical timescale is shorter, where are the timescale values obtained from?
Are they estimated from equation (13)? If this is the case, how do the
theoretical dP/dts are computed?

  > RESPONSE: As noted in the caption of Figure 5, the decay timescale
  > values are obtained from Equation 15, assuming Qstar'=10^7.  We have
  > slightly modified the text for clarity on this point.

@@

REFEREE COMMENT:

v) In the top panels of Figures 4 and 6, the orbital decay model evolution
shows that at the beginning the O-C values increase until, at some point, they
reach a maximum and after that, they start decreasing. What is the physical
explanation for that behavior? If the orbital period is diminishing, wouldn't
it be expected to see only a decreasing in the O-C values? Why the O-C values
would increase at these first epochs?

  > RESPONSE: To make Figure 4 and 6 (now 7), we start with the observed
  > transit and occultation mid-times, as a function of epoch number.
  > We then fit a line to these data.  If the data were generated by a
  > constant period orbit (with gaussian uncertainties) the residuals to
  > such a fit would show no correlation with epoch number.  However if
  > the data were generated by a planet with a constant period
  > derivative (with gaussian uncertainties), the residuals to such a
  > fit would always be a parabola.  This is because the linear fit to
  > the data will have contributions from the beginning, middle, and end
  > of the transit timeseries. Thus the best-fit linear ephemeris (to
  > data that actually have a decreasing period) has a best-fit period
  > somewhere between the minimum and maximum period. At the earliest
  > epochs, the best-fit period is an underestimate, so you see the
  > increase in O-C. At the latest epochs, the best-fit period is an
  > overestimate, so you see the decrease.

@@

REFEREE COMMENT:

ii) In the caption of Figure 7: "with standard deviation (σpredicted)" -->
"with standard deviation (σpre-TESS)"

  > RESPONSE: Corrected.

********************************************************************************
OTHER MINOR CHANGES:

@@

* Added the following coauthors: Fei Dai, Natalia Guerrero, Bill Wohler,
  Douglas Caldwell.

* Minor edits in text throughout, as indicated with the \added{},
  \deleted{}, and \replaced{}{} commands.

* Minor tweaks to figures.

* Added acknowledgement to D. Ragozzine.

* Updated Figure 7 to use physical prior, and to show subset of samples
  with k2 < 0.75.

### 2019-04-11 paper/ms.tex
aj resubmit
aj accepted

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  \replaced{The residuals are the observed times minus the calculated
  times assuming a constant period. The darkest points correspond to
  the most precise data.}{The vertical axis shows the observed times
  minus the calculated times assuming a constant period for transits
  ({\it top}) and occultations ({\it bottom}). In the upper panel,
  darker points correspond to more precise data.} The constant-period
  model (gray line) is a poor description of the data.  Models with a
  decreasing period (blue) or an eccentric, precessing orbit (orange)
  provide	better fits\deleted{ to the data}.  The red square
  represents the combination of all the TESS data and is for display
  purposes only.  The models were fitted to all of the individual
  transit times.
  \label{fig:times}
	}
\end{figure*}

@@

If instead the observed timing variation is just a small portion of an
apsidal precession cycle, then the orbital eccentricity is a few times
$10^{-3}$, and the full precession period is about 27 years.
\citet{ragozzine_probing_2009} calculated apsidal precession periods
for hot Jupiters, finding them to range between about 10 and 100
years. They highlighted that for many hot Jupiters, including WASP-4b,
the theoretical precession rate is dominated by the non-Keplerian
force due to the planet's tidal bulge.  Precession from general
relativity, the planet's rotational bulge, and the star's rotational
and tidal bulges contribute at the 10\% level at most.  Thus, a
measurement of the precession rate can be used to determine the
planet's Love number.  From their Equation 14, the implied Love number
for WASP-4b is
\begin{align}
  k_{2,{\rm p}} &= 1.59^{+0.70}_{-0.47}
    \quad{\rm (wide\ prior),} \label{eq:love_number} \\
  k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5])
\end{align}
where the uncertainties in the semimajor axis and planet radius have been
propagated through our Markov chains, and $\mathcal{U}$ denotes the uniform 
distribution.  For comparison, the Love number
of Jupiter is about 0.55 \citep{wahl_tidal_2016,ni_empirical_2018},
and a uniform density sphere has $k_2 = 1.5$. The uncertainty in
$k_{2,{\rm p}}$ for WASP-4b is large because the eccentricity,
reference time, and ${\rm d}\omega/{\rm d}E$ have strongly correlated
errors, and the \added{four available}\deleted{measured} occultation
times only weakly constrain these parameters
(Figure~\ref{fig:future}).

@@

\acknowledgements
L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
A.~Bailey, J.~Goodman, K.~Patra\added{, D. Ragozzine,} and V.~Van
Eylen, and are grateful to the people who have turned TESS from an
idea into reality.  \added{L.G.B. thanks A.~Bixel and E.~May for
clarifying details concerning the available IMACS lightcurves.}
%
WASP-4 was included on the ``short-cadence'' target list thanks to the
Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
and G011183 respectively). 
%
J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
supporting this work.
%
T.D. acknowledges support from MIT's Kavli Institute as a Kavli
postdoctoral fellow.
%
J.M.D. acknowledges funding from the European Research Council (ERC)
under the European Union's Horizon 2020 research and innovation
programme (grant agreement no. 679633; Exo-Atmos), and support from
the NWO TOP Grant Module 2 (Project Number 614.001.601).
%
J.E.R. was supported by the Harvard Future Faculty Leaders
Postdoctoral fellowship.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} 
  \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  \replaced{The residuals are the observed times minus the calculated
  times assuming a constant period. The darkest points correspond to
  the most precise data.}{The vertical axis shows the observed times
  minus the calculated times assuming a constant period for transits
  ({\it top}) and occultations ({\it bottom}). In the upper panel,
  darker points correspond to more precise data.} The constant-period
  model (gray line) is a poor description of the data.  Models with a
  decreasing period (blue) or an eccentric, precessing orbit (orange)
  provide	better fits\deleted{ to the data}.  The binned TESS
  point (red square) is the weighted average of 18 TESS transits
  and is for display purposes only.
  The models were fitted to all of the individual transit times.
  \label{fig:times}
	}
\end{figure*}

@@

If instead the observed timing variation is just a small portion of an
apsidal precession cycle, then the orbital eccentricity is a few times
$10^{-3}$, and the full precession period is about 27 years.
\citet{ragozzine_probing_2009} calculated apsidal precession periods
for hot Jupiters, finding them to range between about 10 and 100
years. They highlighted that for many hot Jupiters, including WASP-4b,
the theoretical precession rate is dominated by the non-Keplerian
force due to the planet's tidal bulge.  Precession from general
relativity, the planet's rotational bulge, and the star's rotational
and tidal bulges contribute at the 10\% level at most.  Thus, a
measurement of the precession rate can be used to determine the
planet's Love number.  From their Equation 14, the implied Love number
for WASP-4b is
\begin{align}
  k_{2,{\rm p}} &= 1.59^{+0.70}_{-0.47}
    \quad{\rm (wide\ prior),} \label{eq:love_number} \\
  k_{2,{\rm p}} &=  1.20^{+0.20}_{-0.26}
    \quad(k_{2,{\rm p}} \sim \mathcal{U}[0.015, 1.5]),
\end{align}
where the uncertainties in the semimajor axis and planet radius have been
propagated through our Markov chains, and $\mathcal{U}$ denotes the uniform 
distribution.  For comparison, the Love number
of Jupiter is about 0.55 \citep{wahl_tidal_2016,ni_empirical_2018},
and a uniform density sphere has $k_2 = 1.5$. The uncertainty in
$k_{2,{\rm p}}$ for WASP-4b is large because the eccentricity,
reference time, and ${\rm d}\omega/{\rm d}E$ have strongly correlated
errors, and the \added{four available}\deleted{measured} occultation
times only weakly constrain these parameters
(Figure~\ref{fig:future}).

@@

\acknowledgements
L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
A.~Bailey, J.~Goodman, K.~Patra\added{, D. Ragozzine,} and V.~Van
Eylen, and are grateful to the people who have turned TESS from an
idea into reality.  \added{L.G.B. thanks A.~Bixel and E.~May for
clarifying details concerning the available IMACS lightcurves.}
%
WASP-4 was included on the ``short-cadence'' target list thanks to the
Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
and G011183 respectively). 
%
J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
supporting this work.
%
T.D. acknowledges support from MIT's Kavli Institute as a Kavli
postdoctoral fellow.
%
J.M.D. acknowledges funding from the European Research Council (ERC)
under the European Union's Horizon 2020 research and innovation
programme (grant agreement no. 679633; Exo-Atmos), and support from
the NWO TOP Grant Module 2 (Project Number 614.001.601).
%
J.E.R. was supported by the Harvard Future Faculty Leaders
Postdoctoral fellowship.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018},
    Tycho-2 \citep{hog_tycho-2_2000},
    2MASS \citep{skrutskie_two_2006},
    APASS \citep{henden_data_2012},
    WISE \citep{wright_wide-field_2010}.
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} 
  \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}
